A Possibilistic Valid-time Model

### 2012-03-29 trunk/ipmu2012/ipmu.tex
I modified the part about possibilistic variables because I needed the given information to explain the evaluation of sets...
Monday's update!
Some place-saving changes...
Modifications sugested by reviewers + new bibliography.

\usepackage[dvips]{graphicx}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amsfonts}

@@

\mainmatter              % start of the contributions
%
\title{A Possibilistic Valid-time Model}
%
%\titlerunning{A Fuzzy Valid-Time Model}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Jos\'e Enrique Pons \and Christophe Billiet \and Olga Pons Capote \and Guy De Tr\'e}
%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
%Elsa Bertino}
%
\authorrunning{Jos\'e Enrique Pons et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{
Department of Computer Science and Artificial Intelligence \\
Universidad de Granada \\
Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica \\
C/Periodista Daniel Saucedo Aranda s/n \\
E-18071 (Granada-Spain)\\
\email{jpons,opc@decsai.ugr.es}\\ 
%WWW home page:
%\texttt{http://users/\homedir iekeland/web/welcome.html}
%\and
%Universit\'{e} de Paris-Sud,
%Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
%F-91405 Orsay Cedex, France
\and 
Department of Telecommunications and Information Processing
Ghent University
Sint Pietersnieuwstraat 41 B-9000, Ghent, Belgium.
\email{Christophe.Billiet,Guy.De.Tre@ugent.be}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}

\end{abstract}

@@

%%%%%%%%%%%%%%%%
%
% FURTHER RESEARCH WORK
%
%%%%%%%%%%%%%%%
%
\section{Conclusions and further research work}
\label{sec:futher-research}
\input{conclusions}
%

%
\section*{Acknowledgements}
%
The researchers are supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.

>>>>

\usepackage[dvips]{graphicx}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{multirow}

@@

\mainmatter              % start of the contributions
%
\title{A Possibilistic Valid-time Model}
%
%\titlerunning{A Fuzzy Valid-Time Model}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Jos\'e Enrique Pons\inst{1} \and Christophe Billiet\inst{2} \and Olga Pons Capote\inst{1} \and Guy De Tr\'e\inst{2}}
%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
%Elsa Bertino}
%
\authorrunning{Jos\'e Enrique Pons et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{
Department of Computer Science and Artificial Intelligence, Universidad de Granada, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica \\
C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain\\
\email{jpons,opc@decsai.ugr.es}\\ 
%WWW home page:
%\texttt{http://users/\homedir iekeland/web/welcome.html}
%\and
%Universit\'{e} de Paris-Sud,
%Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
%F-91405 Orsay Cedex, France
\and 
Department of Telecommunications and Information Processing, Ghent University \\
Sint-Pietersnieuwstraat 41 B-9000, Ghent, Belgium.
\email{Christophe.Billiet,Guy.De.Tre@ugent.be}
}

\maketitle              % typeset the title of the contribution
\def\JPicScale{0.5}
\begin{abstract}
Information in databases can be imperfect and this imperfection has several forms and causes. In some cases, a single value should be stored, but it is (partially) unknown. The uncertainty about which value to store leads to the aforementioned imperfection. In temporal databases, uncertainty can arise, concerning which temporal notion needs to be stored. Because in temporal databases, temporal notions influence the consistency with which the database models the reality, this uncertainty has a direct impact on the consistency of the model. To represent this temporal uncertainty, previous works have adapted fuzzy sets with conjunctive interpretation, an approach that might prove misleading. This work presents a model that represents the uncertainty using possibility and necessity measures, which are fuzzy sets with disjunctive interpretations.
\end{abstract}

@@

%%%%%%%%%%%%%%%%
%
% FURTHER RESEARCH WORK
%
%%%%%%%%%%%%%%%
%
\section{Conclusions and Further Research}
\label{sec:futher-research}
\input{conclusions}
%

%
\section*{Acknowledgements}
%
Part of the researchers are supported by the grant BES-2009-013805 in the project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in RDBMS}.

### 2012-03-29 trunk/ipmu2012/preliminaries.tex
Modifications sugested by reviewers + new bibliography.

A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge on which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention. This possibility and necessity is noted here as:
\begin{align}
\text{Possibility that $p$ = $T$ (p holds):} \hspace{50pt} & Pos(p) = \pi_p(T) \label{propholdsposs} \\
\text{Necessity that $p$ = $T$ (p holds):} \hspace{50pt} & Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
\end{align}
%Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

This work will deal with ill-known intervals. These are ill-known sets, defined and represented via a start and end point, which will be ill-known values. The elements of the set are the values between the start and end point. A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. The correspondences and transitions between the representations of ill-known sets and between the representation of an ill-known set and an ill-known interval are part of the authors current research.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets\cite{zadeh65} to define a \emph{fuzzy interval}:
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.

@@

The most convenient form of the membership function of a fuzzy number is a triangular function. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. Three real values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$. Here:
\begin{itemize}
\item
$D$ denotes the single value in the core of $M$
\item
$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\item
$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\end{itemize}

The membership function of such a fuzzy number is then given by:

\begin{align}
\mu_M : & \quad \mathbb{R} \rightarrow \left[0,1\right] \nonumber\\
 : & \quad x \rightarrow
\begin{cases}
0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
\frac{1}{a}\left[x-\left(D-a\right)\right] & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
-\frac{1}{b}\left[x-\left(D+b\right)\right] & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
\end{cases}
\end{align}

@@

\begin{eqnarray}
C_1\triangleq\left(\geq,X\right)\\
C_2\triangleq\left(\leq,Y\right).
\end{eqnarray}

To calculate possibility and necessity concerning a conjunction of constraints, the $\min$ operator is used. The possibility and necessity of the set satisfying both constraints is then:

@@

The temporal notions in temporal databases can be classified into 4 types based on their interpretation and modelling purpose. User-defined time has no interpretation, but the other types do:

\begin{itemize}
	\item
	\emph{Transaction time}: The time when the fact is stored in the database.
	\item
	\emph{Valid time}: The time when the fact is true in the modelled reality.
	\item
	\emph{Decision time}: The time when an event was decided to happen. 
\end{itemize}

Database models can then be classified into transaction time \cite{Jensen91}, decision time, valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time) models.

To deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist. %Some fuzzy temporal models assume that the time stored in the database is an interval. The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point. The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.

>>>>

A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention. This possibility and necessity is noted here as:
\begin{align}
\text{Possibility that $p$ = $T$ (p holds):} \hspace{50pt} & Pos(p) = \pi_p(T) \label{propholdsposs} \\
\text{Necessity that $p$ = $T$ (p holds):} \hspace{50pt} & Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
\end{align}
%Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

This work will deal with ill-known intervals. These are ill-known sets, defined and represented via a start and end point, which will be ill-known values. The elements of the set are the values between the start and end point. A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{zadeh65} to define a \emph{fuzzy interval}:
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.

@@

The most convenient form of the membership function of a fuzzy number is a triangular form. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. Three real values, denoted by $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$. Here:
\begin{itemize}
\item
$D$ denotes the single value in the core of $M$
\item
$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\item
$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\end{itemize}

%The membership function of such a fuzzy number is then given by:
%
%\vspace{-10pt}
%
%\begin{align}
%\mu_M : & \quad \mathbb{R} \rightarrow \left[0,1\right] \nonumber\\
% : & \quad x \rightarrow
%\begin{cases}
%0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
%\frac{1}{a}\left[x-\left(D-a\right)\right] & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
%-\frac{1}{b}\left[x-\left(D+b\right)\right] & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
%\end{cases}
%\end{align}

@@

\begin{eqnarray}
C_1 & \triangleq\left(\geq,X\right)\\
C_2 & \triangleq\left(\leq,Y\right)
\end{eqnarray}

To calculate the possibility and necessity concerning a conjunction of constraints, the $\min$ operator can be used. The possibility and necessity of $I$ being included in $\left[X, Y\right]$ are now: %satisfying both constraints is then:

@@

The temporal notions in temporal databases can be classified into four types based on their interpretation and modelling purpose. \emph{User-defined time} has no specific impact on the database consistency, but the other types do:

\begin{itemize}
	\item
	\emph{transaction time}~\cite{Jensen91}: time when the fact is stored in the database.
	\item
	\emph{valid time}~\cite{Snodgrass84}: time when the fact is true in the modelled reality.
	\item
	\emph{decision time}~\cite{Nascimento95}: time when an event was decided to happen. 
\end{itemize}

Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.

%To deal with time points  or intervals  that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist. %Some fuzzy temporal models assume that the time stored in the database is an interval. The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point. The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.

### 2012-03-29 trunk/ipmu2012/proposal.tex
Modifications sugested by reviewers + new bibliography.

\subsubsection{Storage}
\label{subsec:storage}
To store a PVP, the possibility distributions defining the ill-known start and end point are stored. To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a}, \cite{Gal98} is used. Using this representation, it is possible to represent not only fuzzy numbers, but also (fuzzy) constants. To store an ill-known value, four values (FT, F1, F2 and F3) are stored. They are explained in table \ref{table:relational-representation-pvp}. Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database \cite{Medina94gefred.a}.

@@

\subsection{Querying Ill-known Valid-time Intervals}
This subsection discusses a tool for querying. The focus will evidently lie on the querying of valid-time periods. First, the query structure will be defined, followed by the evaluation and ranking methods.

@@

% by a combination of non-temporal query requirements $Q$ is noted $e_{Q}(r) \in \left[0,1\right]$ in this work. The total query evaluation method is now the following.  and Allen relation $AR$
\begin{itemize}
\item
The preferences expressed in $Q$ are evaluated, resulting in a satisfaction degree denoted here $e_{Q}(r)$. The presented model accepts any sound way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$. 
\item
Depending on AR, a specific set of ill-known constraints is considered, which can be found in table 2. The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \eqref{ill-known-pos} respectively \eqref{ill-known-nec} and aggregated using the $\min$ operator.

@@

\subsubsection{Ranking and aggregation}
In order to present the results to the user, a crude ranking method is used: for every record $r$, the sum of $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ gives an evaluation score $e'_{Q^{time}}(r)$ in interval $\left[0,2\right]$. Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural ranking score. This $e'_{Q^{time}}(r)$ is then rescaled to the unit interval, resulting in $e_{Q^{time}}(r)$. The final ranking $e_{final}(r)$ is now given by a convex combination:

@@

\subsection{Example}

%\begin{example} 
\subsubsection{The database}
Consider the example relation given in table \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) describing the approximate period of time during which the car model was sold. The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer. The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model, thus a car model in a certain state.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Sample table for the car models. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
\caption{Example database}
\begin{tabular}{c c c c c c c}
\hline
ID & IID & Segment & Manufacturer & Name & Start & End  \\ [0.5ex]
\hline
001 & 1 & B & Peugeot & 205 & [1985,2,3] & [1997,2,1] \\
002 & 1 & C & Peugeot & 305 & [1977,2,2] & [1989,2,3] \\
003 & 1 & B & Citroen & C2 & [2002,2,2] & [2006,1,1] \\
001 & 2 & B & Peugeot & 206 & [2000,1,2] & [2011,2,1] \\
001 & 3 & B & Peugeot & 207 & [2006,1,1] & [2011,1,1]\\
\hline
\end{tabular}
\vspace{10pt}

@@

\subsubsection{The query}
Consider the following query:
\begin{center}
\emph{The user wants to obtain a list of models from segment B, made by manufacturer Peugeot before the year interval 2001-2005}
\end{center}

@@

\begin{table}[ht]
\caption{Result table and ranking}
\centering
\begin{tabular}{c c c c c c c}
\hline
ID & IID &  $\Pos_{Q^{time}}$ & $\Nec_{Q^{time}}$ & $e_{Q^{time}} (rescaled)$ & $Q$ & $e_{final}$ ($\omega=0.5$) \\ [0.5ex]
\hline
001 & 1 & 1 &  1 & 1 & 1 & 1 \\
002 & 1 & 1 & 1 & 1 & 0.5 & 0.75 \\
003 & 1 & 1 & 0.5 & 0.75 &0 & 0.375\\
001 & 2 & 1 & 0 & 0.5 &1 & 0.75 \\
001 & 3 & 0 & 0 & 0 &1 & 0.5\\
\hline
\end{tabular}
\label{tb:results}
\end{table}
%\end{example}

@@

>>>>

\subsubsection{Storage}
\label{subsec:storage}
To store a PVP, the possibility distributions defining the ill-known start and end point are stored. To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a}, \cite{Gal98} is used. Using this representation, it is possible to represent not only fuzzy numbers, but also (fuzzy) constants. To store an ill-known value, four values (FT, F1, F2 and F3) are stored. They are explained in Table \ref{table:relational-representation-pvp}. Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database \cite{Medina94gefred.a}.

@@

\subsection{Querying Ill-known Valid-time Intervals}
This subsection discusses a tool for querying. Although TSQL~\cite{Jensen1995} is the temporal extension to SQL, the query language proposed is an extension to FSQL~\cite{Gal98} with temporal operators~\cite{Garrido2009}. The focus will evidently lie on the querying of valid-time periods. First, the query structure will be defined, followed by the evaluation and ranking methods.

@@

% by a combination of non-temporal query requirements $Q$ is noted $e_{Q}(r) \in \left[0,1\right]$ in this work. The total query evaluation method is now the following.  and Allen relation $AR$
\begin{itemize}
\item
The preferences expressed in $Q$ are evaluated, resulting in a satisfaction degree denoted here $e_{Q}(r)$. The presented model accepts any sound way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$. 
\item
Depending on AR, a specific set of ill-known constraints is considered, which can be found in Table 2. The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \eqref{ill-known-pos} respectively \eqref{ill-known-nec} and aggregated using the $\min$ operator.

@@

\subsubsection{Ranking and aggregation}
In order to present the results to the user, a crude ranking method is used: for every record $r$, the sum of $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ gives an evaluation score $e'_{Q^{time}}(r)$ in interval $\left[0,2\right]$. Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural ranking score. Some authors ~\cite{Bosc2010a} mentioned before that the possibility and necessity measures result in a total order in the set of events.  This $e'_{Q^{time}}(r)$ is then rescaled to the unit interval, resulting in $e_{Q^{time}}(r)$. The final ranking $e_{final}(r)$ is now given by a convex combination:

@@

%\subsection{Example}

\begin{example}

%\subsubsection{The database}
\paragraph{The database}
Consider the example relation given in Table \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) describing the approximate period of time during which the car model was sold. The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer. The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model, thus a car model in a certain state.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Sample table for the car models. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
\caption{Example database}
\begin{tabular}{c c c c c c c}
\hline
ID & IID & Segment & Manufacturer & Name & Start & End  \\ [0.5ex]
\hline
001 & 1 & B & Peugeot & 205 & [1985,2,3] & [1997,2,1] \\
002 & 1 & C & Peugeot & 305 & [1977,2,2] & [1989,2,3] \\
003 & 1 & B & Citroen & C2 & [2002,2,2] & [2006,1,1] \\
001 & 2 & B & Peugeot & 206 & [2000,1,2] & [2011,2,1] \\
001 & 3 & B & Peugeot & 207 & [2006,1,1] & [2011,1,1]\\
\hline
\end{tabular}
\vspace{10pt}

@@

%\subsubsection{The query}
%\paragraph{The database}
Consider the following query:
\begin{center}
\emph{The user wants to obtain a list of models from segment B, made by manufacturer Peugeot before the year interval 2001-2005}
\end{center}

@@

\begin{table}[ht]
\caption{Result table and ranking}
\centering
\begin{tabular}{c c c c c c c}
\hline
ID & IID &  $\Pos_{Q^{time}}$ & $\Nec_{Q^{time}}$ & $e_{Q^{time}} (rescaled)$ & $Q$ & $e_{final}$ ($\omega=0.5$) \\ [0.5ex]
\hline
001 & 1 & 1 &  1 & 1 & 1 & 1 \\
002 & 1 & 1 & 1 & 1 & 0.5 & 0.75 \\
003 & 1 & 1 & 0.5 & 0.75 &0 & 0.375\\
001 & 2 & 1 & 0 & 0.5 &1 & 0.75 \\
001 & 3 & 0 & 0 & 0 &1 & 0.5\\
\hline
\end{tabular}
\label{tb:results}
\end{table}

@@

\end{example}
